---
title: "<center> <h1>Activity 1</h1> </center>"
author: "Geraldo B. Padilla F."
output:
  word_document: default
  html_document: default
---

# Introduction

In this assignment, we will explore some relationship between variables in a data set referred to Uber and Lyft rideshares in Boston, MA. The data set is a random sample of n= 2.500 observations from 693.071 observations in the original database, available at <https://www.kaggle.com/brllrb/uber-and-lyft-dataset-boston-ma>

```{r, echo=FALSE, warning=FALSE, message=FALSE}
urlfile<-'https://raw.githubusercontent.com/lebebr01/psqf_6243/main/data/rideshare_small.csv'
ride<-read.csv(url(urlfile))
library("easypackages")
paq <- c('psqf6246stat6516','e1071','lubridate', 'mosaic', 'ggformula', 'paramtest', 'pwr', 'ggplot2', 'nlme', 'dplyr', 'dbplyer', 'knitr', 'dplyr', 'dbplyr', 'readxl', 'GGally', 'Hmisc', 'corrplot', 'PerformanceAnalytics', 'statthink')
libraries(paq)
theme_set(theme_statthinking())
```

Specifically, we have the following guiding question: does the distance of the ride explain variation in the price of the ride?

1. In this vein, the first thing to do is exploring the distribution of the price of the rides (shape, center, variation, and/or extreme values)

```{r, echo=FALSE, warning=FALSE, message=FALSE}
list(rsdata)
head(rsdata)
str(rsdata)
```


```{r, echo=FALSE, message=FALSE}
summary(rsdata)
df_stats(~ price, data = rsdata, mean, median, sd, skewness, kurtosis, quantile(probs = c(0.1, 0.5, 0.9)))
```

```{r}

```


```{r, echo=FALSE, eval=FALSE}
gf_histogram(~price, data = rsdata, color = 'black', fill = 'yellow', bins = 30) %>% 
  gf_boxploth(380 ~ price, data = rsdata, fill = "skyblue", width = 12) %>%
  gf_labs(x = "Total price of the ride (in $)", y = 'Frequency')

gf_density(~price, data = rsdata) %>% 
   gf_labs(x = "Total price of the ride (in $)")

()
gf_histogram(~distance, data = rsdata)
```


```{r}
rsdatacor <- rsdata[,-1]
rsdatacor2 <- round(cor(rsdatacor),2)
```


```{r}
corrplot(rsdatacor2, method="number", type="upper")
chart.Correlation(rsdatacor2, histogram = F, pch = 19)
```

```{r}
rsdatacor %>%
  group_by(month)%>%
  mean(rsdata$prices)
```

