---
title: "Apa Table templates"
author: "Geraldo B. Padilla F."
date: "9/14/2021"
output: html_document
---

*Description*: This package creates Word files (.doc files) containing APA style tables for several types of analyses. Using this package minimizes transcription errors and reduces the number commands needed by the user.

```{r}
library('apa')
library('apaTables')
```

1. apa.1way.table

Creates a table of means and standard deviations for a 1-way ANOVA design in APA style.

**Usage**

```{r, eval=FALSE}
apa.1way.table(
iv,
dv,
data,
filename = NA,
table.number = NA,
show.conf.interval = FALSE,
landscape = FALSE
)
```

**Arguments**

`iv`                    Name of independent variable column in data frame

`dv`                   Name of dependent variable column in data frame

`data`                 Project data frame name

`filename`             (optional) Output filename document filename (must end in .rtf or .doc only)

`table.number`         Integer to use in table number output line

`show.conf.interval`  (TRUE/FALSE) Display confidence intervals in table

`landscape`           (TRUE/FALSE) Make RTF file landscape

