---
title: "Second exercise"
author: "Geraldo B. Padilla F."
date: "9/8/2021"
output:
  html_document: default
  word_document: default
---

HEAD
```{r}
rm(list = ls())    # Remove all objects (start with clean slate) 
opar = par()       # Save default graphics parameters as opar
setwd("/RCode")    # Set the working directory
getwd()            # Get/display name of working directory
options(show.signif.stars=FALSE) # Show no stars for significance tests
options(width=72, digits=5, scipen=2) # Control printed output 
ooptions = options(width=72, digits=5, scipen=2) # Save print options
```

Randomize.r, Section 3.9, pages 59-60

```{r}
trtmt = c(1,1,1,2,2,2)  # Create column trtmt = (1,1,1,2,2,2)
ranno = runif(length(trtmt))  # Create column of 6 unif(0,1) RVs
design = data.frame(trtmt,ranno)  # Create data.frame "design""
design  # Display the data.frame design
order(ranno)  # Display positions of ordered RVs, for illustration only
design = design[order(ranno),]  # Sort rows by RVs, save
design  # Display design with rows sorted, for illustration only
design$EU = c(1:6)  # Add col EU = (1,2,3,4,5,6) to design
design  # Display the results of the randomization
```

Soap experiment, Tables 3.10 and 3.11, pages 60 and 63
```{r}
Soap <- c(1,1,1,1,2,2,2,2,3,3,3,3)
Cube <- c(1,2,3,4,5,6,7,8,9,10,11,12)
Prewt <- c(13.14, 13.17, 13.17, 13.17, 13.03, 13.18, 13.12, 13.19, 13.14, 13.19, 13.06, 13.00)
Postwt <- c(13.44, 13.27, 13.31, 12.77, 10.40, 10.57, 10.71, 10.04, 11.28, 11.16, 10.80, 11.18)
WtLoss <- c(-0.30, -0.10, -0.14, 0.40, 2.63, 2.61, 2.41, 3.15, 1.86, 2.03, 2.26, 1.82)
soap.data <- data.frame(Soap, Cube, Prewt, Postwt, WtLoss)
```


# Table 3.10, p60
# Read the data into the data.frame "soap.data"

```{r}
soap.data = read.table("C:/Users/Lenovo/OneDrive - University of Iowa/PhD portfolio/5. Github projects/ISM/First-example/soap.txt", header = TRUE)

head(soap.data, 5)  # Display first 5 lines of soap.data
# Add factor variable fSoap to soap.data frame for later ANOVA
soap.data$fSoap = factor(soap.data$Soap)


design$trtmt
trtmt
design[, 1]
design[, "trtmt"]
```

We could use `trtmt = rep(c(1, 2), each = 3)` to produce 6 treatments, 3 times each.


# Plot WtLoss vs Soap, specify axis labels, suppress x-axis.
plot(WtLoss ~ Soap, data=soap.data, 
     ylab = "Weight Loss (grams)", las=1, xaxt = "n")
# Insert x-axis (axis 1) with tick marks from 1 to 3 by 1.
axis(1, at = seq(1, 3, 1)) 

# Table 3.11, p63
summary(soap.data[,c(1,5:6)]) # Summarize data in cols 1, 5, 6
model1 = aov(WtLoss ~ fSoap, data=soap.data)
anova(model1)
# install.packages("lsmeans")
library(lsmeans)
lsmeans(model1, "fSoap")
=======
1. Exercise 1 using R

Load packages

```{r, warning=FALSE, message=FALSE}
library("easypackages")
paq <- c('psqf6246stat6516','e1071','lubridate', 'mosaic', 'ggformula', 'paramtest', 'pwr', 'ggplot2', 'nlme', 'dplyr', 'dbplyer', 'knitr', 'dplyr', 'dbplyr')
libraries(paq)
```

Create some variables and a database. Then, run some descriptive statistics to explore the database.
We can use different commands to create data depending on the type of variables we need. For instance:

*`runif`: random rational numbers. By default it yields values between 0 a 1, but we can increase the range of expected values. e.g. `runif(100, min=0, max=2)`

*`sample`: random integer numbers, e.g. `sample(18: 65, 100, replace = T)`

*`rnorm`: random numbers clustered around a mean and sd, e.g. `rnorm(100, mean = 1500, sd= 100)`

Let us use `sample` to create some variables.

```{r}
age <- sample(250: 650, 150, replace = T)
income <- sample(250: 1500, 6000, replace = T)
SEG <- sample(250: 1, 4, replace = T)
data1 <- data.frame(age, income, SEG)
```

If we want to enhance the reproducibility of our work, we can set a seed into the randomization. Thus, we will get always the same random numbers.

```{r}
set.seed(1234)
age <- sample(18: 65, 250, replace = T)

set.seed(1234)
income <- sample(750: 1500, 250, replace = T)

set.seed(1234)
SEG <- sample(1: 4, 250, replace = T)

data <- data.frame(age, income, SEG)
```

Now, visualize the database

```{r}
View(data) #you will see the rows and columns that make up your database
summary(data) #you will see some descriptive statistics of the variables in your database 
str(data) #you will see the structure of your database (variables, scale and some values)
length(SEG) # the number of values in the variable
```

Data can also be observed based on groups

```{r}
seg1 = group_by(data1, SEG)
summarise(seg1, mean(age))
```


After exploring the database as a unit, we can see the distribution of the variables.

```{r, echo=FALSE}
gf_histogram(~income, data = data, bins = 50) %>%
  gf_labs(x = 'Incomes')

gf_boxplot(~income, data = data, fill = 'skyblue')%>%
  gf_labs(x = 'Incomes')

gf_histogram(~income, data = data, bins = 50)%>%
  gf_boxplot(12 ~ income, data = data, fill = 'blue', width= 1)%>%
  gf_labs(x= 'Income', y= 'Frequency')

```


the mean is the only value we can use to 'balance' the deviations.


histogram(income)
View(income)



gf_histogram(~age, income)
density(age)
```




2. Varying N (total sample size) and Cohen's d

2.1 Run the function below as it is

```{r}

```


```{r}
power_one_source(Nmin = 50, # min sample size 
                 Nmax = 200, # max sample size
                 Nincrement = 50,  # increment in sample size
                 dmin = .2,  # min efefct size
                 dmax = .5,  # max effect size
                 dincrement = .1, # increment in effect size
                 n.iter = 1000) # number of iterations 
```

  
# 2) keep the effect size as it is and reduce the sample the effect size

# 3) keep the effect size as it is and increase the sample the effect size

# 4) repeat the  2) and 4) but now fix the sample size to its orginal values and manipulate the effect size

# 5) How about trying the 3 groups function? What happends to power now?

power_one_source_3groups(Nmin = 50, # min sample size 
                         Nmax = 200, # max sample size
                         Nincrement = 50,  # increment in sample size
                         dmin = .2,  # min efefct size
                         dmax = .5,  # max effect size
                         dincrement = .1, # increment in effect size
                         n.iter = 1000) # number of iterations  
>>>>>>> 5a3b56387c14c03fd311a03ad1fdeb1330be1f77
