---
title: "Second exercise"
author: "Geraldo B. Padilla F."
date: "9/8/2021"
output: html_document
---

1. Exercise 1 using R

Load packages

```{r, warning=FALSE, message=FALSE}
library("easypackages")
paq <- c('psqf6246stat6516','e1071','lubridate', 'mosaic', 'ggformula', 'paramtest', 'pwr', 'ggplot2', 'nlme', 'dplyr', 'dbplyer', 'knitr', 'dplyr', 'dbplyr')
libraries(paq)
```


```{r}
age <- sample(100: 650, 150, replace = T)
income <- 
gf_histogram(age)
```




2. Varying N (total sample size) and Cohen's d

2.1 Run the function below as it is

```{r}

```


```{r}
power_one_source(Nmin = 50, # min sample size 
                 Nmax = 200, # max sample size
                 Nincrement = 50,  # increment in sample size
                 dmin = .2,  # min efefct size
                 dmax = .5,  # max effect size
                 dincrement = .1, # increment in effect size
                 n.iter = 1000) # number of iterations 
```

  
# 2) keep the effect size as it is and reduce the sample the effect size

# 3) keep the effect size as it is and increase the sample the effect size

# 4) repeat the  2) and 4) but now fix the sample size to its orginal values and manipulate the effect size

# 5) How about trying the 3 groups function? What happends to power now?

power_one_source_3groups(Nmin = 50, # min sample size 
                         Nmax = 200, # max sample size
                         Nincrement = 50,  # increment in sample size
                         dmin = .2,  # min efefct size
                         dmax = .5,  # max effect size
                         dincrement = .1, # increment in effect size
                         n.iter = 1000) # number of iterations  
