---
title: "Intermediate Statistical Methods"
author: "Students"
date: "9/30/2021"
output: html_document
subtitle: Activity 2
---

**Introduction**

This activity explores the relationship between variables in a data set referred to Uber and Lyft rideshares in Boston, MA. The data set is a random sample of 2.500 observations from 693.071 cases in the original database, available at <https://www.kaggle.com/brllrb/uber-and-lyft-dataset-boston-ma>

```{r, echo=FALSE, warning=FALSE, message=FALSE}
urlfile<-'https://raw.githubusercontent.com/lebebr01/psqf_6243/main/data/rideshare_small.csv'
ride<-read.csv(url(urlfile))
library("easypackages")
paq <- c('psqf6246stat6516','e1071','lubridate', 'mosaic', 'ggformula', 'paramtest', 'pwr', 'ggplot2', 'nlme', 'dplyr', 'dbplyer', 'knitr', 'dplyr', 'dbplyr', 'readxl', 'GGally', 'Hmisc', 'corrplot', 'PerformanceAnalytics', 'statthink')
libraries(paq)
theme_set(theme_statthinking())
```

**Guiding question**

Does the distance of the ride explain variation in the price of the ride?

1. This assignment uses the same data from the first activity. Take a few minutes to reacquaint yourself with this analysis, including the linear regression estimates obtained.

```{r}
ride_lm <- lm(price ~ distance, data = ride)
coef(ride_lm)
```

The intercept is 10.44 and the slope is 2.89. The estimated price values will be:

$$
\hat{price} = 10.439 + 2.889 distance
$$


2. Was the intercept from the linear regression fitted in activity 1 using the raw data as is interpretable in the context of the data? Rephrasing slightly, could the intercept term be made more interpretable? If so, how?

Table 1 shows the main descriptive statistics for the two variables included into the simple linear regression 

```{r, message=FALSE}
st1 <- df_stats(~ price, data = ride, mean, min, max, median, sd, skewness, quantile(probs = c(0.25, 0.75)))
st2 <- df_stats(~ distance, data = ride, mean, min, max, median, sd, skewness, quantile(probs = c(0.25, 0.75)))
st <- rbind(st1, st2)
head(st)
```

Based on Table 

3. Fit a modified linear regression that performs a centering. That is, center the distance in some fashion. Options could include, minimum, mean, median, maximum, etc centering. What changed for the estimates obtained from the model?

```{r}
ride <- ride %>%
  mutate(distance_min = distance - min(distance))
ride_lm_min <- lm(price ~ distance_min, data = ride)
coef(ride_lm_min)
coef(ride_lm)
summary(ride_lm_min)$r.square
summary(ride_lm_min)$sigma
```
Applying a minimum-centering transformation to the distance, the y-intercept becomes slightly more interpretable because now there is a value equals zero in the predictor variable. As the structural relationship between the two variables has not changed, the slope still being the same.

4. Which model, the centered or the uncentered, do you feel has a stronger justification for its usage? Be as specific as possible in your rationale.
Use whichever model you feel is the best from #3, then extract from software or compute the standard errors for the estimated regression coefficients. Interpret these standard errors in the context of the problem.

the minimum-centering option gives a more intuitive interpretation.


5. Compute confidence intervals for the two regression coefficients. Justify your choice for level of confidence and interpret the confidence interval.

6. Finally, perform hypothesis testing for the two regression coefficients. In this, set up the null/alternative hypotheses and then interpret the statistical results to provide a statistical conclusion. More specifically, what does the hypothesis test suggest about the null and alternative hypotheses?

The first part talked about the statistical conclusions, now compare/contrast this with practical conclusions. More specifically, would these results be useful in practice? Be as specific as possible in your response.