---
title: "Linear Regression"
author: "Geraldo B. Padilla F."
date: "05-09-2021"
output: html_document
---
**Notes**

1. Exercises

```{r cars}
summary(cars)
```

2. Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

```{r, results='hide', echo=FALSE}
library(tidyverse)
library(ggformula)

colleges <- read_csv(file = "https://raw.githubusercontent.com/lebebr01/statthink/master/data-raw/College-scorecard-clean.csv", 
  guess_max = 10000)
```

- To see the first 6 rows:

```{r}
head(colleges)
```

We can also include an interactive version for viewing the book on the web using the DT package:

DT::datatable(colleges)

To visualize information, we can use Histograms:

```{r}
gf_histogram(~ adm_rate, data = colleges)
```

Symbol ~ helps to identify an exact column in the data object.

We can use colors to the bins:
```{r}
gf_histogram(~ adm_rate, data = colleges, color = 'black', fill = 'yellow')
```

There are, however, some institutions that are quite selective, only admitting fewer than 25% of the students who apply.

We can adjust the width of the bins:
```{r}
gf_histogram(~ adm_rate, data = colleges, color = 'black', fill = 'yellow', bins = 10)
```

or the number of bins:
```{r}
gf_histogram(~ adm_rate, data = colleges, color = 'black', fill = 'yellow', binwidth = .01)

```

Some customizations can be specified using the gf_labs() function.
```{r}
gf_histogram(~adm_rate, data = colleges, color= 'black', fill = 'yellow', bins = 25) %>%
  gf_labs(
    title = "Distribution of admission rates for 2,019 institutions of higher education",
    subtitle = 'Data Source: U.S. Department of Education College Scorecard',
    x = "Admission Rate",
    y = "Frequency",
    caption = "Source: Exercise"
  ) %>%
  gf_theme(theme_classic())
```


```{r}
The theme, theme_statthinking() is included in the statthink library, a supplemental package to the text that can be installed and loaded with the following commands:

install.packages("remotes")
remotes::install_github("lebebr01/statthink")
remotes::install_github('lebebr01/statthink')
library(statthink)  
Attaching package: 'statthink'

- We can then change the theme in a similar manner to how we changed the theme before:

gf_histogram(~ adm_rate, data = colleges, color = 'black', bins = 25) %>%
  gf_labs(
    x = 'Admission Rate',
    y = 'Frequency',
    title = 'Distribution of admission rates for 2,000 institutions of higher education.',
    subtitle = 'Data Source: U.S. Department of Education College Scorecard'
    ) %>%
  gf_theme(theme_statthinking())

- To set a different theme as the default, we will use the theme_set() function and call our theme_statthinking() within this function:

theme_set(theme_statthinking())

- Now when we create a plot, it will automatically use the statthinking theme without having to specify this in the gf_theme() function:

gf_histogram(~ adm_rate, data = colleges, color = 'black', bins = 25) %>%
  gf_labs(
    x = 'Admission Rate',
    y = 'Frequency',
    title = 'Distribution of admission rates for 2,000 institutions of higher education.',
    subtitle = 'Data Source: U.S. Department of Education College Scorecard'
    )

# Density plots

This plot usually highlights similar distributional features as the histogram, but the visualization does not have the same dependency on the specification of bins:

gf_density(~ adm_rate, data = colleges)

In fact, colleges that admit around 75% of their applicants have the highest probability density.

gf_density(~ adm_rate, data = colleges, color = 'black', fill = 'cyan') %>%
  gf_labs(
    x = 'Admission Rate',
    y = 'Probability density',
    title = 'Distribution of admission rates for 2,019 institutions of higher education.',
    subtitle = 'Data Source: U.S. Department of Education College Scorecard'
    )
# Chapter 3

# Load packages
library(tidyverse)
library(ggformula)
library(mosaic)
library(statthink)

3.1 Summarizing Attributes

df_stats(~ adm_rate, data = colleges, median)

df_stats(~ adm_rate, data = colleges, mean) 

- The mean is the only value we can use to “balance” the deviations. However, the mean is “pulled toward the tail"

3.3 Numerically Summarizing Variation

- Range is affected by extreme values

df_stats(~ adm_rate, data = colleges, min, max)

- Provide both percentiles separately

colleges %>%
df_stats(~ adm_rate, quantile(0.10), quantile(.90))

- Provide both percentiles in a single quantile() call

colleges %>%
  df_stats(~ adm_rate, quantile(c(0.1, 0.9)))
  
- Interquartile Range (IQR)

colleges %>%
  df_stats(~ adm_rate, quantile(c(0.25, 0.75)))
  
colleges %>%
  df_stats(~ adm_rate, min, quantile(c(0.25, 0.75)), max)
  
Although our sample of 2,019 institutions of higher education have wildy varying admissions rates (from 0% to 100%), the middle half of those institutions have admissions rates between 55% and 84%. We also note that the 25% of institutions with the lowest admissions rate range from 0% to 55%, while the 25% of institutions with the highest admissions rate range from only 84% to 100%

gf_boxploth(0 ~ adm_rate, data = colleges, fill = "skyblue")  %>%
  gf_labs(x = "Admission rate")

gf_histogram(~ adm_rate, data = colleges, bins = 30) %>%
  gf_boxploth(160 ~ adm_rate, data = colleges, fill = "skyblue", width = 10) %>%
  gf_labs(x = "Admission rate", y = "Frequency")

3.3.4 Empirical Cumulative Density  

gf_ecdf(~ adm_rate, data = colleges) %>%
  gf_labs(x = "Admission rate", y = 'Cumulative proportion')

- an admission rate of 0.71 is associated with a cumulative proportion of 0.50; one-half of the admission rates in the distribution are at or below the value of 0.71.

3.3.5 Variance and Standard Deviation

colleges %>%
  df_stats(~ adm_rate, var, sd)

- In general, the standard deviation is more useful for describing the variation in a sample because it is in the same metric as the data. On average, the observations in the distribution are .211 units from the mean value of .682.

- In general, most observations in a distribution fall within one standard deviation of the mean. So, for our example, most institutions have an admission rate that is between 47.1% and 89.3%.

3.4 Summarizing Categorical Attributes

- Compute category counts

tally(~region, data= colleges)

tally(~region, data = colleges) / 2019

tally(~region, data = colleges, format = "proportion")

3.5 Advanced Extension: Computing Your Own Measure of Variation





1. First, we must load the libraries.

```{r, echo=FALSE, message=FALSE, warning=FALSE}

library(tidyverse)
library(ggformula)
library(mosaic)
library(rsample)
library(statthink)

# Set theme for plots
theme_set(theme_statthinking())

us_weather <- mutate(us_weather, snow_factor = factor(snow), 
                     snow_numeric = ifelse(snow == 'Yes', 1, 0))

count(us_weather, location)
```


```{r}
gf_point(drybulbtemp_min ~ dewpoint_avg, data = us_weather, size = 3, alpha = .2) %>%
  gf_smooth(method = 'lm', size = 1) %>%
  gf_labs(x = "Average daily dew point",
          y = "Minimum daily temperature (in F)")
```

```{r}

cor(drybulbtemp_min ~ dewpoint_avg, data = us_weather, 
    use = 'complete.obs')
```


```{r}
temp_reg <- lm(drybulbtemp_min ~ dewpoint_avg, 
               data = us_weather)
coef(temp_reg)
```

You may notice that the predicted value of minimum temperature increases by 0.90 degrees Fahrenheit for every one degree increase in average daily dew point, often referred to as the linear slope.

This depicts what the y-intercept is, the y-intercept represents the average minimum temperature for an average daily dew point of 0.

More generally, the y-intercept is the average value of the outcome when all of the attributes included in the linear regression are 0.

```{r}
gf_point(drybulbtemp_min ~ dewpoint_avg, data = us_weather, size = 3, alpha = .2) %>%
  gf_smooth(method = 'lm', size = 1) %>%
  gf_vline(xintercept = 0) %>%
  gf_hline(yintercept = 4.83) %>%
  gf_labs(x = "Average daily dew point",
          y = "Minimum daily temperature (in F)")
```


```{r}
sea_temp <- lm(drybulbtemp_min ~ sealevelpressure_avg, data = us_weather)
coef(sea_temp)
```

Therefore, the linear relationship depicted above is extrapolated outside the range of the data (ie, decreased by 29.02 sea level units) which increases the temperature by 610.871 degrees Fahrenheit units.

2. Mean center sea level pressure

First, mean centering the x attribute can often be a way to make the y-intercept more interpretable

```{r}
gf_point(drybulbtemp_min ~ I(sealevelpressure_avg - mean(sealevelpressure_avg, na.rm = TRUE)), 
         data = us_weather, size = 3, alpha = .2) %>%
  gf_smooth(method = 'lm', size = 1) %>%
  gf_labs(x = "Average daily sea level pressure",
          y = "Minimum daily temperature (in F)")
```

```{r}
sealevel_reg_centered <- lm(drybulbtemp_min ~ I(sealevelpressure_avg - mean(sealevelpressure_avg, na.rm = TRUE)), 
         data = us_weather)
coef(sealevel_reg_centered)

df_stats(~sealevelpressure_avg, mean)
```

The difference, is now the y-intercept is much smaller and would represent the mean minimum temperature when the mean-centered sea level pressure is 0 (ie., a value of 30.04 in the original metric).

The value for the y-intercept also makes more intuitive sense, which can aid in model interpretation.

```{r}
sealevel_reg_min <- lm(drybulbtemp_min ~ I(sealevelpressure_avg - min(sealevelpressure_avg, na.rm = TRUE)), 
         data = us_weather)
coef(sealevel_reg_min)

sealevel_reg_max <- lm(drybulbtemp_min ~ I(sealevelpressure_avg - max(sealevelpressure_avg, na.rm = TRUE)), 
         data = us_weather)
coef(sealevel_reg_max)
```

