---
title: "Pieces of code"
author: "Geraldo B. Padilla F."
date: "11/24/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```




```{r}
round(3.835794, digits=2)
round(3.835794-1.026245, digits = 2)
round(3.835794-1.494944, digits = 2)
round(3.835794-1.533106, digits = 2)
```


```{r}
lsmeans(m2, ~ gni_class)
emmeans(m2, ~ gni_class)
```






```{r}
pairwise.t.test(dat$fertility_rate, dat$gni_class, p.adjust = 'none')$p.value
```



```{r}
lm <- lm(fertility_rate ~ gni_class)
broom::tidy(lm)
```

```{r}
lm1 <- lm(fertility_rate ~ 1 + gni_class)
broom::tidy(lm1)
```

```{r}
dat = dat %>%
  mutate(gni_Low_Middle = ifelse(gni_class == 'Low/Middle', 1, 0),
         gni_Upper_Middle = ifelse(gni_class == 'Upper/Middle', 1, 0),
         gni_Upper = ifelse(gni_class == 'Upper', 1, 0))
```

```{r}
m3 = lm(fertility_rate ~ gni_Low_Middle + gni_Upper_Middle + gni_Upper, data = dat)
summary(m1)
summary(m3)
```

```{r}
dat <- dat %>%
  mutate(outfield = ifelse(position == 'outfield', 1, 0),
         infield = ifelse(position == 'infield', 1, 0),
         catcher = ifelse(position == 'catcher', 1, 0))
```



table(gni_class)







```{r}
dat2 %>%
  group_by(gni_dummy) %>%
  summarise(Mean = mean(fertility_rate))

```
22

```{r}
dat$gni2 = factor(dat$gni_class)
dat$gni2 = fct_recode(dat$gni2,
                       '1' = 'Low',
                       '2' = 'Low/Middle',
                       '3' = 'Upper/Middle',
                       '4' = 'Upper')
```

```{r}
dat$gni3 = factor(dat$gni_class)
```


```{r}
dat$d=as.numeric(dat$gni3)
str(dat)
```



str(dat)


table(gni2)

```{r}
(lm(fertility_rate ~ gni2))
```

table(high_gni)


str(dat)

dat$gni3 = as.numeric(factor(dat$gni2))



dat$d <- NULL

```{r}
dat  = dat %>%
  mutate(gni_dummy = ifelse(gni2==1, 1, 0))
table(dat$gni_dummy)
```

```{r}
dat <- dat %>%
  mutate(midwest = ifelse(us_region == 'midwest', 1, 0),
         northeast = ifelse(us_region == 'northeast', 1, 0),
         west = ifelse(us_region == 'west', 1, 0),
         south = ifelse(us_region == 'south', 1, 0),
         outside_US = ifelse(us_region == 'other', 1, 0))

career <- career %>%
  mutate(league_dummy = ifelse(lgID == 'NL', 1, 0))
```




career <- career %>%
  mutate(league_dummy = ifelse(lgID == 'NL', 1, 0))

```{r}
gni_bin <- gni_class
x <- c("Low", "Low/Middle", "Upper", "Upper/Middle")
as.numeric(labels(x, levels = unique(x)))
```

table(gni_class)

```{r}
enaho17_m1$dat <- recode(gni_class$dat, "1:2=0; 4:6=2; 7=3; 8=4")
table(enaho17_m1$region2)
```

```{r}
mutate(dat,
  gni2 = 0 ,
```


```{r}
dat %>%
  mutate(LM = )
```

